Left-linear rules in adhesive categories

### 2024-02-12 icalp/fusione.tex
abstract and intro

\todo[inline]{
This fails for left-linear rules where, as observed
in~\cite{baldan2017domains}, for graph rewriting, the possibility of
merging parts of the state leads to a sort of disjunctive causality,
i.e., a rewriting step, say $r_c$ can be enabled by two different
rewriting steps $r_a$ and $r_b$, in a way that in a derivation
$r_a r_b r_c$ steps $r_b$ and $r_c$ are independent since the presence
of $r_a$ in the past is sufficient for enabling $r_a$, while in
$r_b r_c r_a$ , the steps $r_b$ and $r_c$ are not independent. }

\todo[inline]{
DA RIUSARE\\
Firstly, Church-Rosser theorem does not work out of the box. In
particular, sequential independence between rewriting steps~\cite{}
expressed in terms of the existence of a so-called independence pair
is a canonical and well-understood notion of independence in the
linear case. \todo{P: Too technical? Forse qui e' piu' esplicativo un
  esempio} Whenever two steps are sequential independent
\begin{itemize}
\item they can be switched and 
\item the result of the switch is uniquely determined.
\end{itemize}

Here we observe that, in general, both property can fail: sequential
independent steps could be not switchable and different independent
pairs could exist leading to different switches of the same rewriting
steps.  A refined notion of independence needs to be considered to ensure
that independent rewriting steps can be switched, which is based in the concept
of fillers. The problem was identified
in~\cite{baldan2011adhesivity}, where a subclass of adhesive
categories was identified where independence as induced by a filler
coincides with sequential independence (as induced by an independence
pair). Here it is extended to $\mathcal{M}$-adhesive categories.
}

>>>>

% \todo[inline]{
% This fails for left-linear rules where, as observed
% in~\cite{baldan2017domains}, for graph rewriting, the possibility of
% merging parts of the state leads to a sort of disjunctive causality,
% i.e., a rewriting step, say $r_c$ can be enabled by two different
% rewriting steps $r_a$ and $r_b$, in a way that in a derivation
% $r_a r_b r_c$ steps $r_b$ and $r_c$ are independent since the presence
% of $r_a$ in the past is sufficient for enabling $r_a$, while in
% $r_b r_c r_a$ , the steps $r_b$ and $r_c$ are not independent.
% }

% \todo[inline]{
% DA RIUSARE\\
% Firstly, Church-Rosser theorem does not work out of the box. In
% particular, sequential independence between rewriting steps~\cite{}
% expressed in terms of the existence of a so-called independence pair
% is a canonical and well-understood notion of independence in the
% linear case. \todo{P: Too technical? Forse qui e' piu' esplicativo un
%   esempio} Whenever two steps are sequential independent
% \begin{itemize}
% \item they can be switched and 
% \item the result of the switch is uniquely determined.
% \end{itemize}
% Here we observe that, in general, both property can fail: sequential
% independent steps could be not switchable and different independent
% pairs could exist leading to different switches of the same rewriting
% steps.  A refined notion of independence needs to be considered to ensure
% that independent rewriting steps can be switched, which is based in the concept
% of fillers. The problem was identified
% in~\cite{baldan2011adhesivity}, where a subclass of adhesive
% categories was identified where independence as induced by a filler
% coincides with sequential independence (as induced by an independence
% pair). Here it is extended to $\mathcal{M}$-adhesive categories.
% }
